Applying Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems

### 2015-08-25 Docs/Paper/LaTex/template.tex
Alessandro Review Fixed

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[twocolumn]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{booktabs}
\usepackage{tabu,xcolor,colortbl}
\usepackage{rotating}
\usepackage{amsmath,array}    
\usepackage[utf8]{inputenc} 
\DeclareUnicodeCharacter{FB01}{fi} %Caracter especial
\usepackage{graphicx} %Imagens
\usepackage{listings} % Codes
\usepackage{tabularx} % in the preamble
\usepackage{mathtools} % Serve para as equacoes
\usepackage{cite}

@@

\newcommand{\head}[1]{\textnormal{\textbf{#1}}}
\renewcommand{\lstlistingname}{Algorithm}

\usepackage{color}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{Gray}{gray}{0.93}
\definecolor{DarkGray}{gray}{0.70}

@@

There is no single method available for solving all optimization problems efficiently~\cite{Rao2009}. The most well-known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\ref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors or matrixes that describe the constraints
\begin{equation}
\label{linear-programming-problem}
  minf^t x \: such that  = 
  \begin{cases}
    A.x \leq b, \\ 
    Aeq.x = beq, \\ 
    x \geq 0.
  \end{cases}
\end{equation}

@@

%----------------------------------------------
\subsection{Solving Optimization Problems with Vz}
\label{Optimization-with-Vz}
%----------------------------------------------

An SMT solver is a tool that verifies formula satisfiability based on a set of background theories (e.g., linear arithmetic, bitvectors, and arrays). Z3 is built on top of a Davis–Putnam–Logemann–Loveland solver together with auxiliary algorithms for solving quantifiers formulas~\cite{Bjorner2014}. vZ is one of the tools implemented in the SMT solver Z3; its base function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.  vZ includes an incremental version of  \textcolor{red}{Maximum Resiliency (MaxRes) ~\cite{Federica2008} to Maximum Satisfiability (MaxSAT)~\cite{NarodytskaN}} and a Simplex to solve numbers without defined patterns. In Vz, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives~\cite{Bjorner2015}. Vz provides functions that extend Z3 for solving optimization problems: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

@@

There is also a way to specify critical regions, which is a block of code that is guaranteed to be execute by a single thread at a time. To create a critical region   the directive ``\#pragma omp critical'' should be used.

@@

\begin{itemize}
\item{\textbf{ESBMC:} This strategy uses only one core to solve the partitioning problem, as described in Section~\ref{Verification-Algorithm-using-ESBMC}.
\item{\textbf{ESBMC-MC:} This uses a sequential strategy $\left\{1,2,\ldots, H_{max}\right\}$; there is no parallel controller in this approach, as described in Section~\ref{Multi-core-ESBMC-with-OpenMP}.}
\item{\textbf{ESBMC-PS:} This approach is similar to ESBMC MC; however, it uses a more optimized way to launch the ESBMC instances, as described Section~\ref{Multi-core-ESBMC-with-OpenMP-using-workers}.}
\item{\textbf{ESBMC-PB:} This strategy uses a parallel binary search, as described in Section~\ref{Multi-core-ESBMC-with-OpenMP-using-Binary-Search}.}}
\end{itemize}

There is no single tool for solving all benchmarks efficiently. In particular, Vz solves $4$ out of $7$ benchmarks, but it does not find the optimal solution for the RC$6$ benchmark; \textcolor{blue}{Nao entendi nada desta frase... the slowest result compared with ILP was Patricia that was 3.33 times faster, vZ returned two TO (time outs) in Fuzzy and Mars.}

@@

Table~\ref{experimental-results-matlab-vz-esbmc} shows the experimental results using vZ, Matlab (ILP and GA), and ESBMC (ESBMC-MC, ESBMC-PS, ESBMC-PB) tools.

\begin{table}[h]
\caption {Experimental results of the simple benchmarks}
\small
\begin{tabular}[c]{m{1.3cm}m{1.20cm}|m{0.85cm}|m{0.85cm}|m{0.85cm}|m{0.85cm}}
  \toprule[1.5pt]
  \head{\begin{turn}{-90}  \end{turn}} &
  \head{\begin{turn}{-90}  \end{turn}} &
  \head{\begin{turn}{-90}CRC32\end{turn}} &
  \head{\begin{turn}{-90}Patricia\end{turn}} &
  \head{\begin{turn}{-90}Dijkstra\end{turn}} &
  \head{\begin{turn}{-90}Clustering\end{turn}} \\

  \midrule

\verb| | & \verb|Nodes| & \verb|25| & \verb|21| & \verb|26| & \verb|150| \\
\verb| | &\verb|Edges| & \verb|32| & \verb|48| & \verb|69| & \verb|331| \\
\verb| | &\verb|S0| & \verb|20| & \verb|10| & \verb|20| & \verb|50| \\
\bottomrule[1.5pt]

%EXACT Solution
\rowcolor{DarkGray}
\verb|Exact| & \verb|Hp| & \verb|15| & \verb|47| & \verb|31| & \verb|241| \\
\rowcolor{DarkGray}
\verb|Solution| & \verb|Sp| & \verb|19| & \verb|4| & \verb|19| & \verb|46| \\

%vZ
\verb|vZ| & \verb|T(s)| & \verb|0.2| & \verb|0.3| & \verb|0.6| & \verb|109|\\
\verb|| & \verb|Hp| & \verb|15| & \verb|47| & \verb|31| & \verb|241| \\
\hline

%ILP
\rowcolor{Gray}
\verb|ILP| & \verb|T(s)| & \verb|2| & \verb|1| & \verb|2| & \verb|649|\\
\rowcolor{Gray}
\verb| | & \verb|Hp| & \verb|15| & \verb|47| & \verb|31| & \verb|241|\\
\hline

%GA
\verb|GA| & \verb|T(s)| & \verb|7| & \verb|7| & \verb|9| & \verb|340|\\
\verb| | & \verb|E%| & \verb|13.3| & \verb|0| & \verb|29| & \verb|1.7|\\
\hline

%ESBMC
\rowcolor{Gray}
\verb|ESBMC| & \verb|T(s)| & \verb|31| & \verb|362| & \verb|292| & \verb|3010|\\
\rowcolor{Gray}
\verb|| & \verb|Hp| & \verb|15| & \verb|47| & \verb|31| & \verb|241|\\
\hline

%MultiCore ESBMC
\verb|ESBMC| & \verb|T(s)| & \verb|2| & \verb|6| & \verb|7| & \verb|1615|\\
\verb|MC| & \verb|Hp| & \verb|15| & \verb|47| & \verb|31| & \verb|241|\\
\hline

%ESBMC Parallel sequential
\rowcolor{Gray}
\verb|ESBMC| & \verb|T(s)| & \verb|4| & \verb|10| & \verb|11| & \verb|2238|\\
\rowcolor{Gray}
\verb|PS| & \verb|Hp| & \verb|15| & \verb|47| & \verb|31| & \verb|241|\\
\hline

%ESBMC Parallel Binary
\verb|ESBMC| & \verb|T(s)| & \verb|6| & \verb|5| & \verb|6| & \verb|207|\\
\verb|PB| & \verb|Hp| & \verb|15| & \verb|47| & \verb|38| & \verb|241|\\
\bottomrule[1.5pt]

\end{tabular}
\label{experimental-results-matlab-vz-esbmc}
\end{table}

Table~\ref{Experimental-results-of-the-complex-benchmarks} is an extension of Table~\ref{experimental-results-matlab-vz-esbmc}, where complex benchmarks are used. In particular, Table~\ref{Experimental-results-of-the-complex-benchmarks} shows problems with a higher number of nodes and edges; this increase directly influences the processing time of all tools. RC$6$ produced time out for all versions of ESBMC, vZ e GA did not produce the correct answer, and ILP solves correctly all benchmarks, except one. Fuzzy produced time-outs and memory-outs in all tools. ILP was the only one to solve Mars, while GA solved incorrectly all benchmarks.

\begin{table}[h]
\caption{Experimental results of the complex benchmarks}
\small
\begin{tabular}[c]{m{1.3cm}m{1.25cm}|m{1.25cm}|m{1.25cm}|m{1.25cm}}
  \toprule[1.5pt]
  \head{\begin{turn}{-90}  \end{turn}} &
  \head{\begin{turn}{-90}  \end{turn}} &
  \head{\begin{turn}{-90}RC6\end{turn}} &
  \head{\begin{turn}{-90}Fuzzy\end{turn}} &
  \head{\begin{turn}{-90}Mars\end{turn}} \\

  \midrule

\verb| | & \verb|Nodes| & \verb|329| & \verb|261|  & \verb|417|\\
\verb| | &\verb|Edges| &  \verb|448| & \verb|442|  & \verb|600|\\
\verb| | &\verb|S0| & \verb|600| & \verb|4578|  & \verb|300|\\
\bottomrule[1.5pt]

%EXACT Solution
\rowcolor{DarkGray}
\verb|Exact| & \verb|Hp| & \verb|692|  & \verb|13820| & \verb|876|\\
\rowcolor{DarkGray}
\verb|Solution| & \verb|Sp| & \verb|533|  & \verb|4231| & \verb|297|\\

%vZ
\verb|vZ| & \verb|T(s)| & \verb|112|  & \verb|TO| & \verb|TO|\\
\verb|| & \verb|Hp| & \verb|241|  & \verb|-| & \verb|-|\\
\hline

%ILP
\rowcolor{Gray}
\verb|ILP| & \verb|T(s)| & \verb|1806|  & \verb|TO| & \verb|5.42|\\
\rowcolor{Gray}
\verb| | & \verb|Hp| & \verb|692|  & \verb|-| & \verb|876|\\
\hline

%GA
\verb|GA| & \verb|T(s)| & \verb|2050|  & \verb|1.37| & \verb|5000|\\
\verb| | & \verb|E%| &  \verb|-6.5|  & \verb|-37.6| & \verb|-27.5|\\
\hline

%ESBMC
\rowcolor{Gray}
\verb|ESBMC| & \verb|T(s)| &  \verb|TO|  & \verb|MO| & \verb|MO|\\
\rowcolor{Gray}
\verb|| & \verb|Hp| & \verb|-|  & \verb|-| & \verb|-|\\
\hline

%MultiCore ESBMC
\verb|ESBMC| & \verb|T(s)|  & \verb|TO|  & \verb|TO| & \verb|TO|\\
\verb|MC| & \verb|Hp| & \verb|-|  & \verb|-| & \verb|-|\\
\hline

%ESBMC Parallel sequential
\rowcolor{Gray}
\verb|ESBMC| & \verb|T(s)| & \verb|TO|  & \verb|MO| & \verb|MO|\\
\rowcolor{Gray}
\verb|PS| & \verb|Hp| & \verb|-|  & \verb|-| & \verb|-|\\
\hline

%ESBMC Parallel Binary
\verb|ESBMC| & \verb|T(s)| & \verb|TO|  & \verb|MO| & \verb|MO|\\
\verb|PB| & \verb|Hp| & \verb|-|  & \verb|-| & \verb|-|\\
\bottomrule[1.5pt]

\end{tabular}
\label{Experimental-results-of-the-complex-benchmarks}
\end{table}

Clustering described in Table~\ref{experimental-results-matlab-vz-esbmc} seems to be the limit to test the described tools, note that more than $150$ nodes lead to TO and MO. ILP shows robustness and produces results even for a high amount of nodes and edges.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[twocolumn]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{color}
\usepackage{multirow}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{booktabs}
\usepackage{tabu,xcolor,colortbl}
\usepackage{rotating}
\usepackage{amsmath,array}    
\usepackage[utf8]{inputenc} 
\DeclareUnicodeCharacter{FB01}{fi} %Caracter especial
\usepackage{graphicx} %Imagens
\usepackage{listings} % Codes
\usepackage{tabularx} % in the preamble
\usepackage{mathtools} % Serve para as equacoes
\usepackage{cite}

@@

\newcommand{\head}[1]{\textnormal{\textbf{#1}}}
\renewcommand{\lstlistingname}{Algorithm}
\renewcommand{\arraystretch}{1.3}
\usepackage[table]{xcolor}
\usepackage{color}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{Gray}{gray}{0.93}
\definecolor{DarkGray}{gray}{0.70}

@@

There is no single method available for solving all optimization problems efficiently~\cite{Rao2009}. The most well-known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\ref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors or matrixes that describe the constraints
\begin{equation}
\label{linear-programming-problem}
  minf^t x \: such \; that  = 
  \begin{cases}
    A.x \leq b, \\ 
    Aeq.x = beq, \\ 
    x \geq 0.
  \end{cases}
\end{equation}

@@

@@

There is also a way to specify critical regions, which is a block of code that is guaranteed to be execute by a single thread at a time. To create a critical region   the directive ``\#pragma omp critical'' should be used.
%----------------------------------------------
\subsection{Solving Optimization Problems with Vz}
\label{Optimization-with-Vz}
%----------------------------------------------

An SMT solver is a tool that verifies formula satisfiability based on a set of background theories (e.g., linear arithmetic, bitvectors, and arrays). \textcolor{red}{Z3 is built on the a Davis–Putnam–Logemann–Loveland algorithm together with auxiliary algorithms for solving quantifiers formulas~\cite{Bjorner2014}}. vZ is one of the tools implemented in the SMT solver Z3; its base function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.  vZ includes an incremental version of  \textcolor{red}{Maximum Resiliency (MaxRes) ~\cite{Federica2008} to Maximum Satisfiability (MaxSAT)~\cite{NarodytskaN}} and a Simplex to solve numbers without defined patterns. In Vz, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives~\cite{Bjorner2015}. Vz provides functions that extend Z3 for solving optimization problems: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

@@

\begin{itemize}
\item{\textbf{ESBMC:} This strategy uses only one core to solve the partitioning problem, as described in Section~\ref{Verification-Algorithm-using-ESBMC}.
\item{\textbf{ESBMC-MC:} This uses a sequential strategy \{1 ,2 ,\:...\:, Hmax\}; there is no parallel controller in this approach, as described in Section~\ref{Multi-core-ESBMC-with-OpenMP}.}
\item{\textbf{ESBMC-PS:} This approach is similar to ESBMC MC; however, it uses a more optimized way to launch the ESBMC instances, as described Section~\ref{Multi-core-ESBMC-with-OpenMP-using-workers}.}
\item{\textbf{ESBMC-PB:} This strategy uses a parallel binary search, as described in Section~\ref{Multi-core-ESBMC-with-OpenMP-using-Binary-Search}.}}
\end{itemize}

\textcolor{red}{There is no single tool for solving all benchmarks efficiently. In particular, Vz solves $4$ out of $7$ benchmarks, but it does not find the optimal solution for the RC$6$ benchmark; in the worst case, vZ was more fast than ILP 3.33 times (Patricia code),vZ returned two TO (time outs) in Fuzzy and Mars.}

@@

Table~\ref{Experimental-results-of-the-complex-benchmarks} shows the experimental results using vZ, Matlab (ILP and GA), and ESBMC (ESBMC\hyp{}MC, ESBMC\hyp{}PS, ESBMC\hyp{}PB) tools.

RC$6$ produced time out for all versions of ESBMC, vZ e GA did not produce the correct answer, and ILP solves correctly all benchmarks, except one. Fuzzy produced time-outs and memory-outs in all tools. ILP was the only one to solve Mars, while GA solved incorrectly all benchmarks.

Clustering described in Table~\ref{Experimental-results-of-the-complex-benchmarks} seems to be the limit to test the described tools, note that more than $150$ nodes lead to TO and MO. ILP shows robustness and produces results even for a high amount of nodes and edges.

\begin{table*}[t]
  \centering
  \caption{Experimental results of the complex benchmarks}  
  \begin{tabular}{*{9}{cc|c|c|c|c|c|c|c}}

 &  & \textbf{CRC32} & \textbf{Patricia} & \textbf{Dijkstra} & \textbf{Clustering} & \textbf{RC6} & \textbf{Fuzzy} & \textbf{Mars} \\[0.15cm]
\cmidrule(r){3-9}

&\textbf{Nodes} &25 &21 &26 &150 &329 &261 &417\\
&\textbf{Edges} &32 &48 &69 &331 &448 &442 &600\\
&\textbf{S0} &20 &10 &20 &50 &600 &4578 &300\\

%EXACT Solution
\bottomrule[1.5pt]
\rowcolors{1}{}{lightgray}

\multirow{2}{*}{\textbf{Exact Solution}}
&  Hp   &  15   &  47   &  31   &  241   &  692    &  13820   &  876 \\
&  Sp   &  19   &  4   &  19   &  46   &  533    &  4231   &  297  \\

%vZ
\bottomrule[1.5pt]
\multirow{2}{*}{\textbf{vZ}}
&  T(s)   &  0.2   &  0.3   &  0.6   &  109   &  112    &  TO   &  TO  \\
&  Hp   &  15   &  47   &  31   &  241   &  241    &  -   &  -  \\
\hline

%ILP
\multirow{2}{*}{\textbf{ILP}}  
&  T(s)   &  2   &  1   &  2   &  649   &  1806    &  TO   &  5.42  \\
&  Hp   &  15   &  47   &  31   &  241   &  692    &  -   &  876  \\
\hline

%GA
\multirow{2}{*}{\textbf{GA}}
&  T(s)   &  7   &  7   &  9   &  340   &  2050    &  1.37   &  5000  \\
&  Error \%   &  13.3   &  0   &  29   &  1.7   &  -6.5    &  -37.6   &  -27.5  \\
\hline

%ESBMC
\multirow{2}{*}{\textbf{ESBMC}}
&  T(s)   &  31   &  362   &  292   &  3010   &  TO    &  MO   &  MO  \\
&  Hp   &  15   &  47   &  31   &  241   &  -    &  -   &  -  \\
\hline

%MultiCore ESBMC
\multirow{2}{*}{\textbf{ESBMC Multicore}}
&  T(s)   &  2   &  6   &  7   &  1615   &  TO    &  TO   &  TO  \\
&  Hp   &  15   &  47   &  31   &  241   &  -    &  -   &  -  \\
\hline

%ESBMC Parallel sequential
\multirow{2}{*}{\textbf{ESBMC Parallel Sequential}}
&  T(s) &  4 &  10   &  11   &  2238   &  TO    &  MO   &  MO  \\
&  Hp   &  15   &  47   &  31   &  241   &  -    &  -   &  -  \\
\hline

%ESBMC Parallel Binary
\multirow{2}{*}{\textbf{ESBMC Parallel Binary}}
&  T(s)   &  6   &  5   &  6   &  207   &  TO    &  MO   &  MO \\
&  Hp   &  15   &  47   &  38   &  241   &  -    &  -   &  -  \\
\bottomrule[1.5pt]
\end{tabular}
\label{Experimental-results-of-the-complex-benchmarks}
\end{table*}
